---
title: ""
params:
  area: ""
date: "`r Sys.Date()`"
output:
  blogdown::html_page:
    toc: true
---

```{r setup, include=FALSE, out.width="100%"}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(reactable)

# Parameters
  # Parameters for state and data
file_data_params <- here::here("data/params.yml")
  # Household Pulse data for state
file_hps <- here::here("data/state.csv")
  # Unemployment data
file_unemployment <- here::here("data/unemployment.csv")
  # Food insufficiency estimates for counties
file_counties <- here::here("data/counties.csv")

#===============================================================================

# Parameters for state and data
data_params <- yaml::read_yaml(file_data_params)

# Household Pulse data for state
hps <- 
  read_csv(file_hps) %>% 
  filter(area_type == "State", variable == "curfoodsuf", code %in% 3:4)

# Unemployment data
unemployment <- 
  read_csv(file_unemployment) %>% 
  filter(area_type == "State" | area == params$area)

# State
state <- 
  unemployment %>% 
  filter(area_type == "State") %>% 
  distinct(area) %>% 
  pull(area)

# Food insufficiency estimates for county
county <- 
  read_csv(file_counties) %>% 
  filter(area == params$area)
```

## Food insufficiency

```{r}
v <- 
  bind_rows(
    county %>% 
      group_by(area, date_end, variable) %>% 
      summarize(
        n = sum(n),
        pct = sum(pct)
      ) %>% 
      ungroup(),
    hps %>% 
      group_by(area, date_end, variable) %>% 
      summarize(
        n = sum(n),
        pct = sum(pct)
      ) %>% 
      ungroup()
  ) %>% 
  mutate(area = factor(area, levels = c(params$area, state)))

v %>% 
  ggplot(aes(date_end, pct, color = area)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end,
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Food insufficiency during the prior 7 days",
    x = "Survey end date",
    y = "Percentage of adults in households with insufficient food",
    color = NULL,
    caption = 
      "Estimated from: Census Bureau, Household Pulse Survey.\nBureau of Labor Statistics, Local Area Unemployment Statistics."
  )
```

<br>

```{r}
v %>% 
  filter(area == params$area) %>% 
  select(date_end, n, pct) %>% 
  arrange(desc(date_end)) %>% 
  mutate(
    date_end = format(date_end, "%Y %b %e") %>% str_replace("  ", " "),
    n = signif(n, digits = 3),
    pct = format(round(pct, digits = 1), nsmall = 1)
  ) %>% 
  select(
    "Survey end date" = date_end, 
    "Estimated percentage of adults reporting food insufficiency" = pct,
    "Estimated number of adults reporting food insufficiency" = n
  ) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list(
        "Estimated percentage of adults reporting food insufficiency" = 
          colDef(format = colFormat(suffix = "%"))
      )
  )

v <- 
  v %>% 
  filter(area == params$area, date_end == max(date_end))
```

In the most recent survey period that ended `r format(v$date_end, "%B %e")`, an estimated `r format(round(v$pct, digits = 1), nsmall = 1)`% of `r params$area` adults, or approximately `r format(signif(v$n, digits = 3), big.mark = ",")` adults, lived in households experiencing food insufficiency.

Because the Household Pulse Survey does not provide data at the county level, we use a model to estimate the county-level food insufficiency numbers. The model uses the food insufficiency rate of `r data_params$state` and the county unemployment rates to estimate food insufficiency for each county.

The state-level food insufficiency data is updated more frequently than the county unemployment rates, and each month's unemployment data comes out later than the food insufficiency data. If unemployment data is not yet available for a month, the model uses the latest month for which data is available. We revise the county-level estimates for such months as new unemployment data becomes available. 

## Unemployment

```{r}
unemployment %>% 
  mutate(area = factor(area, levels = c(params$area, state))) %>% 
  ggplot(aes(date, unemployment_rate, color = area)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = unemployment$date,
    minor_breaks = NULL,
    date_labels = "%b"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Unemployment rate",
    x = NULL,
    y = "Unemployment rate",
    color = NULL,
    caption = 
      "Source: Bureau of Labor Statistics, Local Area Unemployment Statistics."
  )
```

<br>

```{r}
unemployment %>% 
  filter(area == params$area) %>% 
  arrange(desc(date)) %>% 
  mutate(
    date = format(date, "%Y %b"),
    unemployment_rate = format(unemployment_rate, nsmall = 1)
  ) %>% 
  select(
    Month = date,
    "Unemployment rate" = unemployment_rate,
    "Unemployment" = unemployment
  ) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list("Unemployment rate" = colDef(format = colFormat(suffix = "%")))
  )
```


The data for the most recent month is typically "preliminary" and subject to revision when the following month's data is released.

