---
params:
  area: ""
title: ""
date: "Updated: `r Sys.Date()`"
output:
  blogdown::html_page:
    toc: true
---

```{r setup, include=FALSE, out.width="100%"}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)

# Parameters
  # Parameters for state and data
file_data_params <- here::here("data/params.yml")
  # Household Pulse data for state
file_hps <- here::here("data/state.csv")
  # Population data
file_population <- here::here("data/population.csv")
  # Unemployment data
file_unemployment <- here::here("data/unemployment.csv")
  # Food insecurity estimates for counties
file_counties <- here::here("data/counties.csv")
  # Number of bins for maps  
N_BINS <- 5
  # Map colors
MAP_COLORS <- RColorBrewer::brewer.pal(n = N_BINS, name = "YlOrRd")

#===============================================================================

# Parameters for state and data
data_params <- yaml::read_yaml(file_data_params)
# Population data for counties
population <- 
  read_csv(file_population) %>% 
  filter(area_type == "County")
# Household Pulse data for state
hps <- 
  read_csv(file_hps) %>% 
  filter(area_type == "State")
# Last dates for Household Pulse data
hps_date_start <- max(hps$date_start)
hps_date_end<- max(hps$date_end)
# Last date for Household Pulse PUF data
hps_puf_date_end <- 
  hps %>% 
  drop_na(n_error) %>% 
  pull(date_end) %>% 
  max()
# Unemployment data
unemployment <- read_csv(file_unemployment)
# Food insecurity estimates for county
counties <- read_csv(file_counties)
# County boundaries for map
if (is_null(data_params$state_proj)) {
  county_boundaries <-
    ussf::boundaries(geography = "county") %>% 
    filter(STATEFP == data_params$state_fips)
} else {
  county_boundaries <-
    ussf::boundaries(geography = "county") %>% 
    st_transform(crs = data_params$state_proj) %>% 
    filter(STATEFP == data_params$state_fips)
}

#===============================================================================

# Bin variable into quantiles weighted by population
bin_var <- function(var, population) {
  cut(
    var, 
    breaks = 
      Hmisc::wtd.quantile(
        var,
        probs = seq(0, 1, length.out = N_BINS + 1),
        weights = population
      ),
    include.lowest = TRUE
  )
}
# Choropleth map of variable
choropleth <- function(data, var) {
  county_boundaries %>% 
    left_join(data, by = c("GEOID" = "fips")) %>% 
    ggplot(aes(fill = {{var}})) +
    geom_sf(size = 0.1) +
    scale_fill_manual(
      values = MAP_COLORS, 
      labels = scales::label_percent(accuracy = 0.1, scale = 1)
    ) +
    guides(
      fill = 
        guide_colorsteps(show.limits = TRUE, barwidth = 15, barheight = 0.25)
    ) +
    theme_void() +
    theme(legend.position = "bottom")
}
```

## Overview

The data on this site is derived from the U.S. Census Bureau [Household Pulse Survey (HPS)](https://www.census.gov/householdpulsedata) and the U.S. Bureau of Labor Statistics [Local Area Unemployment Statistics (LAUS)](https://www.bls.gov/lau/home.htm). New HPS data is released once a week, and new LAUS data is released once a month. The most recent data used for this report are:

* Household Pulse Survey: `r format(hps_date_start, "%B %e")` - `r format(hps_date_end, "%B %e")`
* Local Area Unemployment Statistics: `r format(max(unemployment$date), "%B")`

The HPS releases data in two stages. For this reason, some time series below run through `r format(hps_puf_date_end, "%B %e")`, while others run through `r format(hps_date_end, "%B %e")`.

Upon the release of new data, this report will be updated. You can sign up with [this form](https://forms.gle/MssCSstca6egwMDF6) to be notified when the report has been updated with new data. All of the data used in this report can be found and downloaded from [here](https://github.com/stanford-datalab/covid-19-ca/tree/master/data). All of the code used to download and process the raw data and to generate this site can be found [here](https://github.com/stanford-datalab/covid-19-ca).

If you have any feedback, comments, or suggestions for improvements to this site, please [email us](mailto:datalab@stanford.edu).

## Food insecurity

```{r}
prior_food_insecurity <- 
  hps %>%
  filter(variable == "prifoodsuf", code %in% 3:4) %>% 
  group_by(date_end) %>% 
  summarize(prior_food_insecurity = sum(pct)) %>% 
  pull(prior_food_insecurity) %>% 
  mean()
v <- 
  hps %>% 
  group_by(date_end) %>% 
  summarize(
    food_insecurity = sum(pct[variable == "curfoodsuf" & code %in% 3:4])
  )
v %>% 
  ggplot(aes(date_end, food_insecurity)) +
  geom_hline(yintercept = prior_food_insecurity, color = "blue") +
  geom_line() +
  geom_point() +
  annotate(
    geom = "text",
    x = min(v$date_end),
    y = prior_food_insecurity,
    label = "Food insecurity prior to Mar 13",
    hjust = 0,
    vjust = 1.5
  ) +
  scale_x_date(
    breaks = v$date_end,
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  labs(
    title = "Food insecurity in adults",
    x = "Survey end date",
    y = "Percentage of adults who are food insecure",
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

The above plot contains the percentage of adults who answered this question:

>  In the __last 7 days__, which of these statements best describes the food eaten in your household?

with one of these two answers:

> Sometimes not enough to eat  
> Often not enough to eat

The HPS was not designed to measure food insecurity in children, but other surveys typically find food insecurity in children to be substantially higher than that for adults.
<br>
<br>

```{r fig.asp=0.8}
v <- 
  hps %>% 
  filter(str_detect(variable, "^foodsufrsn\\d_curfoodsuf_34$"), !is.na(code))
v %>% 
  ggplot(aes(date_end, pct, color = fct_reorder2(response, date_end, pct))) +
    geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Reasons for food insecurity in adults",
    x = "Survey end date",
    y = "",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

The above plot contains the reasons for those who selected "Sometimes not enough to eat" or "Often not enough to eat". Since respondents can choose more than one reason, they sum to greater than 100%.
<br>
<br>

```{r}
v <- 
  counties %>% 
  filter(date_end == max(date_end)) %>% 
  group_by(area, fips, date_end, variable) %>% 
  summarize(food_insecurity = sum(pct)) %>% 
  ungroup() %>% 
  left_join(population %>% select(fips, population_18p), by = "fips") %>% 
  mutate(food_insecurity = bin_var(food_insecurity, population_18p))
v %>% 
  choropleth(food_insecurity) +
  labs(
    title = "Food insecurity in adults",
    subtitle = format(max(counties$date_end), "%B %e, %Y"),
    fill = NULL,
    caption = 
      "Estimated from: Census Bureau, Household Pulse Survey.\nBureau of Labor Statistics, Local Area Unemployment Statistics."
  )
```

These estimates of food insecurity in `r data_params$state`'s counties are based upon a model. The model uses the overall rate of food insecurity for the state, which is updated every week, and the unemployment rate for each county, which is updated every month. The unemployment data for a month comes out later than the food insecurity data. For months where unemployment data not yet available, the model uses the latest month for which data is available.
<br>
<br>

```{r}
v <- 
  hps %>% 
  filter(variable == "curfoodsuf", code %in% 3:4) %>% 
  group_by(date_end) %>% 
  mutate(prop = n / sum(n))
v %>% 
  ggplot(aes(date_end, prop, color = fct_inorder(response))) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, NA)
  ) +
  scale_color_discrete(direction = -1) +
  theme(legend.position = "bottom") +
  labs(
    title = "Type of food insecurity in adults",
    x = "Survey end date",
    y = "Percentage of type of food insecurity",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

<br>
<br>

```{r}
v <- 
  hps %>% 
  filter(variable == "freefood", code == 1)
v %>% 
  ggplot(aes(date_end, pct)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end,
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  labs(
    title = 
      "During the last 7 days, did you or anyone in\nyour household get free groceries or a free meal?",
    x = "Survey end date",
    y = "Percentage of adults in households that received free food",
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

<br>
<br>

```{r fig.asp=0.8}
v <- 
  hps %>% 
  filter(str_detect(variable, "^wherefree"), code == 1)
v %>% 
  ggplot(aes(date_end, pct, color = fct_inorder(response))) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Where did you get free groceries or free meals?",
    x = "Survey end date",
    y = 
      "Percentage who received free food from this source",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

Since respondents can choose more than one source of free groceries or free meals, they sum to greater than 100%.
<br>
<br>

## Unemployment

```{r}
v <- 
  unemployment %>% 
  filter(area_type == "State") %>% 
  arrange(date)
  
v %>% 
  filter(area_type == "State") %>% 
  ggplot(aes(date, unemployment_rate)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date,
    minor_breaks = NULL,
    date_labels = "%b"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  labs(
    title = "Unemployment rate",
    x = NULL,
    y = "Unemployment rate",
    caption = 
      "Source: Bureau of Labor Statistics, Local Area Unemployment Statitics."
  )
```

The data for the most recent month is typically "preliminary" and subject to revision when the following month's data is released.
<br>
<br>

```{r}
v <-
  unemployment %>% 
  filter(area_type == "County", date == max(date)) %>% 
  left_join(population %>% select(fips, population_18p), by = "fips") %>% 
  mutate(unemployment_rate = bin_var(unemployment_rate, population_18p))
v %>% 
  choropleth(unemployment_rate) +
  labs(
    fill = NULL,
    title = "Unemployment rate",
    subtitle = format(unique(v$date), "%B %Y"),
    caption = 
      "Source: Bureau of Labor Statistics, Local Area Unemployment Statistics."
  )
```
