---
params:
  area: ""
title: ""
date: "`r Sys.Date()`"
output:
  blogdown::html_page:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(reactable)
library(rstanarm)
library(sf)

# Parameters
  # Parameters for state and data
file_data_params <- here::here("data/params.yml")
  # Model parameters
file_params_model <- here::here("data/params_model.yml")
  # Population data
file_population <- here::here("data/population.csv")
  # Household Pulse data for state
file_hps <- here::here("data/state.csv")
  # Unemployment data
file_unemployment <- here::here("data/unemployment.csv")
  # Food insufficiency estimates for counties
file_counties <- here::here("data/counties.csv")
  # Unemployment and food insufficiency data from Food Security Supplement
file_unemployment_fi_fss <- 
  here::here("data/unemployment_food_insufficiency_fss.csv")
  # First month for Pulse path
FIRST_PULSE_MONTH <- ymd("2020-05-01")
  # Number of bins for maps  
N_BINS <- 5
  # Map color palette
MAP_COLOR_PALETTE <- "YlOrRd"
  # Map height
MAP_HEIGHT <- 650
  # Age group breaks
AGE_BREAKS <- c(18, seq(29, 79, 10), Inf)
  # Age group labels
AGE_LABELS <- 
  c("18 - 29", str_c(seq(30, 70, 10), " - ", seq(39, 79, 10)), "80+")

#===============================================================================

# Parameters for state and data
data_params <- yaml::read_yaml(file_data_params)

# Parameters for model of child food insufficiency
child_model <- yaml::read_yaml(file_params_model)$child_model

# Population data for state and counties
population <- 
  file_population %>% 
  read_csv(
    col_types = 
      cols(
        area_type = col_character(),
        area = col_character(),
        fips = col_character(),
        population = col_double(),
        population_0_17 = col_double(),
        population_18p = col_double()
      )
  )

# Population of children in state
population_0_17 <- 
  population %>% 
  filter(area_type == "State") %>% 
  pull(population_0_17)

# Population data for counties
population <- 
  population %>% 
  filter(area_type == "County")

# Household Pulse data for state
hps <- 
  file_hps %>% 
  read_csv(
    col_types = 
      cols(
        area_type = col_character(),
        area = col_character(),
        fips = col_character(),
        date_start = col_date(format = ""),
        date_end = col_date(format = ""),
        variable = col_character(),
        code = col_double(),
        response = col_character(),
        n = col_double(),
        n_error = col_double(),
        pct = col_double()
      )
  ) %>% 
  filter(area_type == "State")

# Last dates for Household Pulse data
hps_date_start <- max(hps$date_start)
hps_date_end <- max(hps$date_end)

# Last dates for Household Pulse PUF data
v <- 
  hps %>% 
  drop_na(n_error)
hps_puf_date_start <- max(v$date_start)
hps_puf_date_end <- max(v$date_end)

# Unemployment data
unemployment <- 
  file_unemployment %>% 
  read_csv(
    col_types = 
      cols(
        area_type = col_character(),
        area = col_character(),
        fips = col_character(),
        date = col_date(format = ""),
        status = col_character(),
        labor_force = col_double(),
        employment = col_double(),
        unemployment = col_double(),
        unemployment_rate = col_double()
      )
  )

# Last date with statewide unemployment data
unemployment_state_date <- 
  unemployment %>% 
  filter(area_type == "State") %>% 
  pull(date) %>% 
  max()

# Last date with full county-level unemployment data
n_counties <- 
  unemployment %>% 
  filter(area_type == "County") %>% 
  pull(fips) %>% 
  n_distinct()
unemployment_county_date <- 
  unemployment %>% 
  filter(area_type == "County") %>% 
  group_by(date) %>% 
  filter(n() == n_counties) %>% 
  pull(date) %>% 
  max()

# Food insufficiency estimates for county
counties <- 
  file_counties %>% 
  read_csv(
    col_types = 
      cols(
        area = col_character(),
        fips = col_character(),
        date_start = col_date(format = ""),
        date_end = col_date(format = ""),
        variable = col_character(),
        code = col_double(),
        response = col_character(),
        n = col_double(),
        pct = col_double()
      )
  )

# County boundaries for map
county_boundaries <-
  ussf::boundaries(
    geography = "county",
    resolution = "5m",
    projection = "longlat"
  ) %>% 
  filter(STATEFP == data_params$state_fips)

# State unemployment and food insufficiency data from Food Security Supplement
unemployment_fi_fss <- 
  file_unemployment_fi_fss %>% 
  read_csv(
    col_types = 
      cols(
        area = col_character(),
        fips = col_character(),
        date = col_date(format = ""),
        labor_force = col_double(),
        employment = col_double(),
        unemployment = col_double(),
        unemployment_rate = col_double(),
        food_insufficiency_n = col_double(),
        food_insufficiency_pct = col_double()
      )
  ) %>% 
  filter(area == params$area)

# State food insufficiency from most recent Food Security Supplement
fi_fss_last <- 
  unemployment_fi_fss %>% 
  transmute(
    year = year(date),
    pct = food_insufficiency_pct
  ) %>% 
  filter(year == min(2019, max(year)))

# State unemployment for months of Household Pulse Survey
unemployment_hps <- 
  unemployment %>% 
  filter(
    area == params$area,
    date >= FIRST_PULSE_MONTH
  ) %>% 
  mutate(unemployment_rate = 100 * unemployment / labor_force) %>% 
  select(-area_type, -status)

# State food insufficiency for months of Household Pulse Survey
fi_hps <- 
  hps %>% 
  filter(
    variable == "curfoodsuf",
    str_detect(response, "not enough to eat")
  ) %>% 
  group_by(area, fips, date_start, date_end) %>% 
  summarize(across(c(n, pct), sum, .names = "food_insufficiency_{col}")) %>% 
  ungroup()

# Household Pulse Survey weeks that correspond to unemployment survey weeks
unemployment_fi_dates <- 
  tibble(
    date = unemployment_hps$date,
    date_end = 
      make_date(year = year(date), month = month(date), day = 12L) %>% 
      map(~ min(fi_hps$date_end[. <= fi_hps$date_end])) %>% 
      reduce(c)
  )

# State unemployment and food insufficiency for months of Household Pulse Survey
unemployment_fi_hps <- 
  unemployment_fi_dates %>% 
  left_join(unemployment_hps, by = "date") %>% 
  left_join(fi_hps %>% select(-area, -fips), by = "date_end") %>% 
  relocate(area, fips, date, date_start, date_end)

# State unemployment and food insufficiency from Food Security Supplement and
# for Household Pulse Survey
unemployment_fi_all <- 
  unemployment_fi_fss %>% 
  select(date, unemployment_rate, food_insufficiency_pct) %>% 
  bind_rows(
    unemployment_fi_hps %>% 
      select(date, unemployment_rate, food_insufficiency_pct)
  )

#===============================================================================

# Month of date
month <- function(date) {
  format(date, "%B")
}

# Month and day of date
month_day <- function(date) {
  format(date, "%B %e") %>%
    str_replace("  ", " ")
}

# Choropleth map
choropleth <- function(data, var, population, tooltip, legend_title) {
  
  values <- 
    data %>% 
    pull({{var}})
    
  bins <-
    Hmisc::wtd.quantile(
      values,
      weights = data %>% pull({{population}}),
      probs = seq(0, 1, length.out = N_BINS + 1)
    )
  
  values[near(values, min(bins))] <- min(bins)
  values[near(values, max(bins))] <- max(bins)

  fill_color <- 
    colorBin(
      palette = MAP_COLOR_PALETTE,
      domain = values,
      bins = bins
    )
  
  tooltips <-
    data %>% 
    str_glue_data(tooltip) %>% 
    map(htmltools::HTML)
  
  data %>%
    leaflet(
      height = MAP_HEIGHT,
      options = leafletOptions(minZoom = 3)
    ) %>%
    setMapWidgetStyle(list(background = "white")) %>% 
    addProviderTiles(
      provider = "MapBox",
      options =
        providerTileOptions(
          id = "mapbox.light",
          accessToken = Sys.getenv("MAPBOX_ACCESS_TOKEN")
        )
    ) %>%
    addPolygons(
      color = "black",
      weight = 0.2,
      opacity = 1,
      fillColor = ~ fill_color(values),
      fillOpacity = 1,
      smoothFactor = 0,
      label = tooltips,
      labelOptions =
        labelOptions(
          style = list("font-weight" = "normal", padding = "3px 8px"),
          textsize = "15px",
          direction = "auto"
        ),
      highlightOptions =
        highlightOptions(
          color = "black",
          weight = 1.5,
          bringToFront = TRUE
        )
    ) %>%
    addLegend(
      position = "topright",
      pal = fill_color,
      values = values,
      opacity = 1,
      labFormat = labelFormat(suffix = "%", digits = 1),
      title = legend_title
    )
}
```

## Overview

The data on this site is derived from the U.S. Census Bureau [Household Pulse Survey (HPS)](https://www.census.gov/householdpulsedata) and the U.S. Bureau of Labor Statistics [Local Area Unemployment Statistics (LAUS)](https://www.bls.gov/lau/home.htm). New HPS data is currently released every two weeks, and new LAUS data is released once a month. The most recent data used for this report are:

* Household Pulse Survey: 
  * Summary data: `r month_day(hps_date_start)` - `r month_day(hps_date_end)`
  * Complete data: `r month_day(hps_puf_date_start)` - `r month_day(hps_puf_date_end)`
* Local Area Unemployment Statistics:
  * Statewide data: `r month(unemployment_state_date)`
  * County-level data: `r month(unemployment_county_date)`

This report updates upon the release of new data. Sign up with [this form](https://forms.gle/MssCSstca6egwMDF6) to be notified of updates. All data used in this report can be found and downloaded [here](https://github.com/stanford-datalab/covid-19-ca/tree/master/data). All code used to download and process the raw data and to generate this site can be found [here](https://github.com/stanford-datalab/covid-19-ca).

If you have any feedback, comments, or suggestions for improvements to this site, please [email us](mailto:datalab@stanford.edu).

## Food insufficiency

### Rate over time

The plot below contains the percentage of adults who answered this question:

>  In the __last 7 days__, which of these statements best describes the food eaten in your household?

with one of these two answers:

> Sometimes not enough to eat  
> Often not enough to eat

```{r}
v <- 
  hps %>% 
  group_by(date_end) %>% 
  summarize(
    adult_pct = sum(pct[variable == "curfoodsuf" & code %in% 3:4]),
    adult_n = sum(n[variable == "curfoodsuf" & code %in% 3:4])
  ) %>% 
  mutate(
    child_pct = child_model$intercept + child_model$slope * adult_pct,
    child_n = (child_pct / 100) * population_0_17
  )

v %>% 
  ggplot(aes(date_end, adult_pct)) +
  geom_hline(yintercept = fi_fss_last$pct, color = "blue") +
  geom_line() +
  geom_point() +
  annotate(
    geom = "text",
    x = min(v$date_end),
    y = fi_fss_last$pct,
    label = 
      str_glue("Food insufficiency in Dec {fi_fss_last$year}"),
    hjust = 0,
    vjust = 1.5
  ) +
  scale_x_date(
    breaks = v$date_end,
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  theme(axis.text.x = element_text(size = rel(0.75), angle = -45, hjust = 0)) +
  labs(
    title = "Food insufficiency during the prior 7 days",
    x = "Survey end date",
    y = "Percentage of adults reporting food insufficiency",
    caption = "Sources: Census Bureau, Household Pulse Survey.\nCurrent Population Survey, Food Security Supplement."
  )
```

<br>

```{r}
v %>% 
  arrange(desc(date_end)) %>% 
  mutate(
    date_end = format(date_end, "%Y %b %e") %>% str_replace("  ", " ")
  ) %>% 
  mutate(
    across(ends_with("_pct"), ~ format(round(., digits = 1), nsmall = 1)),
    across(ends_with("_n"), ~ signif(round(.), digits = 3))
  ) %>% 
  select(
    "Survey end date" = date_end, 
    "Percentage of adults reporting food insufficiency" = adult_pct,
    "Number of adults reporting food insufficiency" = adult_n,
    "Estimated percentage of children experiencing food insufficiency" = child_pct,
    "Estimated number of children experiencing food insufficiency" = child_n
  ) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list(
        "Percentage of adults reporting food insufficiency" = 
          colDef(format = colFormat(suffix = "%", digits = 1)),
        "Number of adults reporting food insufficiency" =
          colDef(format = colFormat(digits = 0, separators = TRUE)),
        "Estimated percentage of children experiencing food insufficiency" = 
          colDef(format = colFormat(suffix = "%", digits = 1)),
        "Estimated number of children experiencing food insufficiency" =
          colDef(format = colFormat(digits = 0, separators = TRUE))
      )
  )

v <- 
  v %>% 
  filter(date_end == max(date_end))
```

In the most recent survey period that ended `r month_day(v$date_end)`, an estimated `r format(round(v$adult_pct, digits = 1), nsmall = 1)`% of `r params$area` adults, or approximately `r format(signif(round(v$adult_n), digits = 3), big.mark = ",")` adults, lived in households with food insufficiency.

Using a model based upon data from the Food Security Supplement of the Current Population Survey, we estimate that in the most recent survey period, `r format(round(v$child_pct, digits = 1), nsmall = 1)`% of children, or approximately `r format(signif(round(v$child_n), digits = 3), big.mark = ",")` children, experienced food insufficiency.

### Geography

The map below is for the survey period that ended `r month_day(hps_date_end)`. Hover over a county for its food insufficiency rate.

```{r}
v <- 
  left_join(
    county_boundaries,
    counties %>% 
      filter(date_end == max(date_end)) %>% 
      group_by(area, fips, date_end, variable) %>% 
      summarize(
        adult_pct = sum(pct),
        adult_n = sum(n)
      ) %>% 
      ungroup() %>% 
      left_join(
        population %>% 
          select(fips, population_0_17, population_18p),
        by = "fips"
      ),
    by = c("GEOID" = "fips")
  ) %>% 
  mutate(
    child_pct = child_model$intercept + child_model$slope * adult_pct,
    child_n = (child_pct / 100) * population_0_17
  )

v %>% 
  choropleth(
    var = adult_pct,
    population = population_18p,
    tooltip = 
      "<strong>{area}</strong><br/>Food insufficiency: {format(round(adult_pct, digits = 1), trim = TRUE)}%",
    legend_title = "Food insufficiency<br/>in adults"
  )
```

<br>

```{r}
v %>% 
  st_drop_geometry() %>% 
  mutate(
    across(ends_with("_pct"), ~ format(round(., digits = 1), nsmall = 1)),
    across(ends_with("_n"), ~ signif(round(.), digits = 3))
  ) %>% 
  select(
    "County" = NAME, 
    "Estimated percentage of adults experiencing food insufficiency" = 
      adult_pct,
    "Estimated number of adults experiencing food insufficiency" = 
      adult_n,
    "Estimated percentage of children experiencing food insufficiency" = 
      child_pct,
    "Estimated number of children experiencing food insufficiency" = 
      child_n
  ) %>% 
  arrange(County) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list(
        "Estimated percentage of adults experiencing food insufficiency" = 
          colDef(format = colFormat(suffix = "%", digits = 1)),
        "Estimated number of adults experiencing food insufficiency" =
          colDef(format = colFormat(digits = 0, separators = TRUE)),
        "Estimated percentage of children experiencing food insufficiency" = 
          colDef(format = colFormat(suffix = "%", digits = 1)),
        "Estimated number of children experiencing food insufficiency" =
          colDef(format = colFormat(digits = 0, separators = TRUE))
      )
  )
```

Estimated from: Census Bureau, Household Pulse Survey.  
Bureau of Labor Statistics, Local Area Unemployment Statistics.

Because the Household Pulse Survey does not provide data at the county level, we use a model to estimate the county-level food insufficiency numbers. The model uses the food insufficiency rate of `r data_params$state` and the county unemployment rates to estimate food insufficiency for each county.

The state-level food insufficiency data is updated more frequently than the county unemployment rates, and each month's unemployment data comes out later than the food insufficiency data. If unemployment data is not yet available for a month, the model uses the latest month for which data is available. 

### Demographics

```{r}
v <- 
  hps %>% 
  filter(
    date_end == hps_puf_date_end,
    str_detect(variable, "^tbirth_year")
  ) %>% 
  mutate(
    age = lubridate::year(lubridate::now()) - code,
    age_group = 
      cut(
        age, 
        breaks = AGE_BREAKS,
        labels = AGE_LABELS,
        include.lowest = TRUE
      )
  ) %>% 
  group_by(age_group, variable) %>% 
  summarize(n = sum(n)) %>% 
  ungroup()

overall_rate <- 
  v %>% 
  summarize(
    pct = 
      100 * sum(n[variable == "tbirth_year_curfoodsuf_34"]) / 
      sum(n[variable == "tbirth_year_curfoodsuf_1234"]),
    label = "Overall food insufficiency rate"
  )

v <- 
  v %>% 
  group_by(age_group) %>% 
  summarize(
    pct =
      100 * n[variable == "tbirth_year_curfoodsuf_34"] / 
      n[variable == "tbirth_year_curfoodsuf_1234"]
  )

v %>% 
  ggplot(aes(age_group, pct)) +
  geom_hline(aes(yintercept = pct, color = label), data = overall_rate) +
  geom_col() +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Food insufficiency by age group",
    subtitle = 
      str_glue(
        "For the survey period that ended {month_day(hps_puf_date_end)}"
      ),
    x = "Age group",
    y = "Percentage of age group reporting food insufficiency",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

<br>
<br>

```{r}
v <- 
  hps %>% 
  filter(
    date_end == hps_puf_date_end,
    str_detect(variable, "^egender")
  )

overall_rate <- 
  v %>% 
  summarize(
    pct = 
      100 * sum(n[variable == "egender_curfoodsuf_34"]) / 
      sum(n[variable == "egender_curfoodsuf_1234"]),
    label = "Overall food insufficiency rate"
  )

v <- 
  v %>% 
  group_by(response) %>% 
  summarize(
    pct =
      100 * n[variable == "egender_curfoodsuf_34"] / 
      n[variable == "egender_curfoodsuf_1234"]
  )

v %>% 
  ggplot(aes(response, pct)) +
  geom_hline(aes(yintercept = pct, color = label), data = overall_rate) +
  geom_col() +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Food insufficiency by gender",
    subtitle = 
      str_glue(
        "For the survey period that ended {month_day(hps_puf_date_end)}"
      ),
    x = NULL,
    y = "Percentage of gender reporting food insufficiency",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

<br>
<br>

```{r}
v <- 
  hps %>% 
  filter(
    date_end == hps_puf_date_end,
    str_detect(variable, "^rhispanic_rrace")
  ) %>% 
  mutate(
    response =
      case_when(
        str_detect(response, "^Hispanic") ~ "Hispanic",
        TRUE ~ response
      )
  ) %>% 
  group_by(variable, response) %>% 
  summarize(n = sum(n)) %>% 
  ungroup()

overall_rate <- 
  v %>% 
  summarize(
    pct = 
      100 * sum(n[variable == "rhispanic_rrace_curfoodsuf_34"]) / 
      sum(n[variable == "rhispanic_rrace_curfoodsuf_1234"]),
    label = "Overall food insufficiency rate"
  )

v <- 
  v %>% 
  group_by(response) %>% 
  summarize(
    pct =
      100 * n[variable == "rhispanic_rrace_curfoodsuf_34"] / 
      n[variable == "rhispanic_rrace_curfoodsuf_1234"]
  )

v %>% 
  ggplot(aes(fct_reorder(response, pct), pct)) +
  geom_hline(aes(yintercept = pct, color = label), data = overall_rate) +
  geom_col() +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Food insufficiency by race and ethnicity",
    subtitle = 
      str_glue(
        "For the survey period that ended {month_day(hps_puf_date_end)}"
      ),
    x = NULL,
    y = "Percentage of group reporting food insufficiency",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

In the above, _Hispanic_ refers to individuals of any race. _White_, _Black_, and _Asian_ refer to individuals of that single race. _Other_ refers to individuals of two or more races, or races other the White, Black, or Asian.

### Reasons

```{r fig.asp=0.8}
v <- 
  hps %>% 
  filter(str_detect(variable, "^foodsufrsn\\d_curfoodsuf_34$"), !is.na(code))

v %>% 
  ggplot(aes(date_end, pct, color = fct_reorder2(response, date_end, pct))) +
    geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme(
    axis.text.x = element_text(size = rel(0.75), angle = -45, hjust = 0),
    legend.position = "bottom"
  ) +
  labs(
    title = "Reasons for food insufficiency in adults",
    x = "Survey end date",
    y = "",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

The above plot contains the reasons respondents gave for selecting "Sometimes not enough to eat" or "Often not enough to eat". Respondents can choose more than one reason, so the percentages sum to greater than 100%.

### Severity

```{r}
v <- 
  hps %>% 
  filter(variable == "curfoodsuf", code %in% 3:4) %>% 
  group_by(date_end) %>% 
  mutate(prop = n / sum(n))

v %>% 
  ggplot(aes(date_end, prop, color = fct_inorder(response))) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, NA)
  ) +
  scale_color_discrete(direction = -1) +
  theme(
    axis.text.x = element_text(size = rel(0.75), angle = -45, hjust = 0),
    legend.position = "bottom"
  ) +
  labs(
    title = "Type of food insufficiency in adults",
    x = "Survey end date",
    y = "Percentage of type of food insufficiency",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

### Free groceries or meals

```{r}
v <- 
  hps %>% 
  filter(variable == "freefood", code == 1)

v %>% 
  ggplot(aes(date_end, pct)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end,
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  theme(axis.text.x = element_text(size = rel(0.75), angle = -45, hjust = 0)) +
  labs(
    title = 
      "During the last 7 days, did you or anyone in\nyour household get free groceries or a free meal?",
    x = "Survey end date",
    y = "Percentage of adults in households that received free food",
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

<br>
<br>

```{r fig.asp=0.8}
v <- 
  hps %>% 
  filter(str_detect(variable, "^wherefree"), code == 1) %>% 
  mutate(response = str_wrap(response, width = 50))

v %>% 
  ggplot(aes(date_end, pct, color = fct_reorder2(response, date_end, pct))) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date_end %>% unique() %>% sort(),
    minor_breaks = NULL,
    date_labels = "%b %e"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  guides(color = guide_legend(ncol = 2)) +
  theme(
    axis.text.x = element_text(size = rel(0.75), angle = -45, hjust = 0),
    legend.position = "bottom"
  ) +
  labs(
    title = "Where did you get free groceries or free meals?",
    x = "Survey end date",
    y = 
      "Percentage who received free food from this source",
    color = NULL,
    caption = "Source: Census Bureau, Household Pulse Survey."
  )
```

Since respondents can choose more than one source of free groceries or free meals, the sources sum to greater than 100%.
<br>
<br>

## Unemployment

### Rate over time

```{r}
v <- 
  unemployment %>% 
  filter(area_type == "State") %>% 
  arrange(date)
  
v %>% 
  filter(area_type == "State") %>% 
  ggplot(aes(date, unemployment_rate)) +
  geom_line() +
  geom_point() +
  scale_x_date(
    breaks = v$date,
    minor_breaks = NULL,
    date_labels = "%b"
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(2),
    labels = scales::label_percent(accuracy = 1, scale = 1),
    limits = c(0, NA)
  ) +
  labs(
    title = "Unemployment rate",
    x = NULL,
    y = "Unemployment rate",
    caption = 
      "Source: Bureau of Labor Statistics, Local Area Unemployment Statistics."
  )
```

<br>

```{r}
v %>% 
  arrange(desc(date)) %>% 
  mutate(
    date = format(date, "%Y %b"),
    unemployment_rate = format(unemployment_rate, nsmall = 1)
  ) %>% 
  select(
    Month = date,
    "Unemployment rate" = unemployment_rate,
    "Unemployment" = unemployment
  ) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list(
        "Unemployment rate" = 
          colDef(format = colFormat(suffix = "%", digits = 1))
      )
  )
```

The data for the most recent month is typically "preliminary" and subject to revision when the following month's data is released.

### Geography

The map below is for `r month(unemployment_county_date)`. Hover over counties for rates.

```{r}
v <- 
  left_join(
    county_boundaries,
    unemployment %>% 
      filter(area_type == "County", date == unemployment_county_date) %>% 
      left_join(population %>% select(fips, population_18p), by = "fips"),
    by = c("GEOID" = "fips")
  )

v %>% 
  choropleth(
    var = unemployment_rate,
    population = population_18p,
    tooltip = 
      "<strong>{area}</strong><br/>Unemployment rate: {format(unemployment_rate, trim = TRUE, nsmall = 1)}%",
    legend_title = "Unemployment rate"
  )
```

<br>

```{r}
v %>%
  st_drop_geometry() %>% 
  mutate(unemployment_rate = format(unemployment_rate, nsmall = 1)) %>% 
  select(
    "County" = NAME,
    "Unemployment rate" = unemployment_rate,
    "Unemployment" = unemployment
  ) %>% 
  arrange(County) %>% 
  reactable(
    showSortable = TRUE,
    defaultColDef = 
      colDef(align = "left", format = colFormat(separators = TRUE)),
    columns = 
      list(
        "Unemployment rate" = 
          colDef(format = colFormat(suffix = "%", digits = 1))
      )
  )
```

Source: Bureau of Labor Statistics, Local Area Unemployment Statistics.

## Unemployment and food insufficiency

```{r}
model <- 
  stan_glm(
    formula = food_insufficiency_pct ~ unemployment_rate, 
    data = unemployment_fi_fss,
    refresh = 0
  )
```

```{r, message=FALSE}
labels_fss <- 
  unemployment_fi_fss %>% 
  filter(
    date == min(date) | 
      unemployment == max(unemployment) |
      food_insufficiency_pct == max(food_insufficiency_pct) |
      date == max(date)
  ) %>% 
  mutate(
    label = format(date, "%Y %b"), 
    hjust = c(0.1, -0.1, -0.1, 0.1),
    vjust = c(1.8, 0.5, 0.5, 1.8)
  )

labels_hps <- 
  unemployment_fi_hps %>% 
  filter(date == min(date) | date == max(date)) %>% 
  mutate(
    label = format(date, "%Y %b"),
    hjust = c(1.1, 1.1),
    vjust = c(0.5, 0.5)
  )

ggplot(mapping = aes(unemployment_rate, food_insufficiency_pct)) +
  geom_path(aes(alpha = date), data = unemployment_fi_fss) +
  geom_point(aes(alpha = date), data = unemployment_fi_fss) +
  geom_text(
    aes(label = label, hjust = hjust, vjust = vjust),
    data = labels_fss
  ) +
  geom_path(data = unemployment_fi_hps) +
  geom_point(data = unemployment_fi_hps) +
  geom_text(
    aes(label = label, hjust = hjust, vjust = vjust),
    data = labels_hps
  ) +
  geom_abline(
    slope = coef(model)[["unemployment_rate"]],
    intercept = coef(model)[["(Intercept)"]],
    color = "blue"
  ) +
  expand_limits(y = 3) +
  coord_fixed() +
  scale_x_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(1),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  theme(legend.position = "none") +
  labs(
    title = 
      str_glue("{params$area} unemployment and food insufficiency over time"),
    x = "Unemployment rate",
    y = "Percentage of adults reporting food insufficiency",
    caption = 
      str_c(
        "Sources: Census Bureau, Household Pulse Survey.\n",
        "Current Population Survey, Food Security Supplement.\n",
        "Bureau of Labor Statistics, Local Area Unemployment Statistics."
      )
  )
```

We modeled food insufficiency on the state unemployment rate to understand how the current situation relates to the past. The blue line shows our model. 

The relationship between food insufficiency and unemployment that held between 2001 and 2019 appears to have changed during the pandemic. Since the outbreak of the pandemic, food insufficiency is much higher than it was in the past for the same levels of unemployment.

